---
title: "615 FEMA mapping"
author: "Jingyi Niu"
date: "2020/11/12"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyverse)
library(rgdal)
library(magrittr)
library(lubridate)
library(shiny)
library(rgeos)
library(sp)
library(leaflet)
library(leaflet.extras)
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this report, I'll show how I produce map with hurricane information from 2009 to 2018.I find two datasets from FEMA website below:

https://www.fema.gov/openfema-data-page/disaster-declarations-summaries-v2 ; https://www.fema.gov/openfema-data-page/public-assistance-funded-projects-details-v1

Summaries data is a summarized dataset describing all federally declared disasters. This dataset lists all official FEMA Disaster Declarations, beginning with the first disaster declaration in 1953 and features all three disaster declaration types: major disaster, emergency, and fire management assistance. The dataset includes declared recovery programs and geographic areas (county not available before 1964; Fire Management records are considered partial due to historical nature of the dataset).
Detail data is called "the Public Assistance Funded Projects Details dataset", it contains obligated (financial obligation to grantee) Public Assistance projects, lists public assistance recipients designated as applicants in the data, and a list of every funded, individual project, called project worksheets. Open projects still under pre-obligation processing are not represented.


```{r}
#### Filter out data from 2009-2018

FEMAsummary <- read.csv("C:/Users/dell/Desktop/DisasterDeclarationsSummaries.csv")
FEMAdetail <- read.csv("C:/Users/dell/Desktop/PublicAssistanceFundedProjectsDetails.csv")
#typeof(FEMAdetail$declarationDate)
h_summary <- FEMAsummary %>% filter(incidentType == "Hurricane", year(as.Date(declarationDate))>=2009 , year(as.Date(declarationDate))<=2018)
h_detail <- FEMAdetail %>% filter(incidentType =="Hurricane",  year(as.Date(declarationDate))>=2009 , year(as.Date(declarationDate))<=2018)
##write.csv(h_detail,file = "detail.csv")
##write.csv(h_summary,file = "summaries2.csv")


#### Read data
smas <- read.csv('E:/bos mssp/MA615/615 FEMA new assignment/615-FEMAmapping/hurricane/data/summaries.csv', stringsAsFactors = FALSE) 
detail <- read.csv('E:/bos mssp/MA615/615 FEMA new assignment/615-FEMAmapping/hurricane/data/detail.csv', stringsAsFactors = FALSE) 
```


## Prepare data


```{r}
# choose data we need, to reduce amount of data and string 

smas <- smas %>% 
    select(
        state_name = state, fipsStateCode, fipsCountyCode, fyDeclared, 
        declarationTitle, disasterNumber, designatedArea
    ) %>% 
    filter(!(fipsCountyCode == 0)) %>% 
    filter(!(fipsStateCode %in% c(78, 60)))
detail <- detail %>% 
    mutate(fyDeclared = as.numeric(substr(declarationDate, 1, 4))) %>% 
    select(fyDeclared, disasterNumber, 
           fipsStateCode = stateNumberCode, 
           fipsCountyCode = countyCode, federalShareObligated, projectAmount) %>% 
    group_by(fyDeclared, disasterNumber, fipsStateCode, fipsCountyCode) %>% 
    summarise(
        federalShareObligated = sum(federalShareObligated, na.rm = TRUE), 
        projectAmount = sum(projectAmount, na.rm = TRUE)
    )


# connect data for map 
plot.data <- smas %>%
    left_join(detail, by = c('fipsStateCode', 'fipsCountyCode', 'fyDeclared', 'disasterNumber'))

# prepare data for shiny 
years <- sort(unique(plot.data$fyDeclared))
# declarationTitles <- sort(unique(plot.data$declarationTitle))
# designatedAreas <- sort(unique(plot.data$designatedArea))

# Prepare for America map data
us_map <- readOGR('E:/bos mssp/MA615/615 FEMA new assignment/615-FEMAmapping/hurricane FEMA/hurricane/data/gz_2010_us_050_00_20m/gz_2010_us_050_00_20m.shp')
us_map@data <- us_map@data %>% 
    mutate(
        fipsStateCode = as.numeric(STATE),
        fipsCountyCode = as.numeric(COUNTY)
    )

# smas %>% select(fipsStateCode) %>% filter(!(fipsStateCode %in% us_map@data$fipsStateCode)) %>% unique()
# smas %>% select(fipsCountyCode) %>% filter(!(fipsCountyCode %in% us_map@data$fipsCountyCode)) %>% unique()

```


## Map for shiny

```{r}
source('global.R', encoding = 'UTF-8')

# Define UI for application that draws a histogram
ui <- fluidPage(
    
    # Application title
    titlePanel("Hurricane"),
    
    # Sidebar with a slider input for number of bins 
    sidebarLayout(
        sidebarPanel(
            selectInput('year', 'Choice Year: ', years), 
            uiOutput('hurricane'), 
            uiOutput('hurricane_sta')
        ),
        
        # Show a plot of the generated distribution
        mainPanel(
            leafletOutput("usMapPlot")
        )
    )
)

# Define server logic required to draw a histogram
server <- function(input, output) {
    
    values <- reactiveValues(
        year_data         = NULL, # data filtered by year
        declarationTitleS = NULL, # hurricane data filter,depends on year
        year_state        = NULL, # state data filtered by year
        hurricane_state   = NULL  # satate filter,depends on year and hurricane type
    )
    
    observeEvent(input$year, {
        values$year_data <- plot.data %>% filter(fyDeclared == input$year) 
        values$declarationTitleS <- sort(unique(values$year_data$declarationTitle)) 
        output$hurricane <- renderUI({
            radioButtons(
                inputId = "declarationTitle", 
                label = "Choice Hurricane: ", 
                c('ALL Hurricane', values$declarationTitleS), 
                'ALL Hurricane', inline = FALSE
            )
        })
    })
    
    observeEvent(input$declarationTitle, {
        if (input$declarationTitle != 'ALL Hurricane') {
            values$year_state <- values$year_data %>% 
                filter(declarationTitle == input$declarationTitle) %>% 
                select(state_name, fipsStateCode) %>% unique()
            
            if (length(unique(values$year_state$state_name)) > 1) {
                values$hurricane_state <- sort(unique(values$year_state$state_name))
                output$hurricane_sta <- renderUI({
                    radioButtons(
                        inputId = "STATE_NAME",
                        label = "Choice State Name: ",
                        c('ALL STA', values$hurricane_state), 
                        'ALL STA', inline = TRUE
                    )
                })
            } else {
                values$hurricane_state <- c(sort(unique(values$year_state$state_name)))
                output$hurricane_sta <- renderUI({
                    radioButtons(
                        inputId = "STATE_NAME",
                        label = "State Name: ",
                        values$hurricane_state, inline = TRUE
                    )
                })
            }
        }
    })
    
    observeEvent(input$declarationTitle, {
        # All County map
        output$usMapPlot <- renderLeaflet({
            if (input$declarationTitle == 'ALL Hurricane') {
                map <- us_map
                # Prepare for disctintuish state data 
                map_sta_boder <- aggregate(
                    map[, "STATE"], by = list(ID = map@data$STATE), FUN = unique, dissolve = T
                )
                # actual data
                map_data <- values$year_data %>%
                    group_by(state_name, fipsStateCode, fipsCountyCode, designatedArea) %>%
                    summarise(
                        federalShareObligated = sum(federalShareObligated, na.rm = TRUE),
                        projectAmount = sum(projectAmount, na.rm = TRUE)
                    )
                
                map@data <- map@data %>%
                    left_join(map_data, by = c('fipsStateCode', 'fipsCountyCode')) %>%
                    mutate(
                        TYPE = ifelse(is.na(designatedArea), 'No Hurricane', ifelse(
                            is.na(projectAmount) | projectAmount <= 0, 'Hurricane No PA', 'Hurricane PA'))
                    )
                
                map@data$text <- sprintf(
                    "<strong>County: %s</strong><br/>TYPE: %s<br/>",
                    map@data$NAME, map@data$TYPE
                ) %>% lapply(htmltools::HTML)
                
                pal <- colorFactor(
                    palette = c('#998ec3', '#f1a340', '#f7f7f7'),
                    domain = map@data$TYPE
                )
                leaflet(map) %>%
                    setView(-96, 37.8, 4) %>%
                    setMapWidgetStyle(list(background = "#f7f7f7")) %>%
                    addPolygons(
                        color = 'Grey',
                        opacity = 0.2, smoothFactor = 0.3,
                        fillOpacity = 1,
                        weight = 1,
                        fillColor = ~ pal(map@data$TYPE),
                        label = map@data$text
                    ) %>%
                    addPolylines(data = map_sta_boder, color = "DarkGrey", opacity = 1, weight = 0.8)
            }
        })
    })
    
    observeEvent(input$STATE_NAME, {
        # Specific hurricane map
        output$usMapPlot <- renderLeaflet({
            if (input$STATE_NAME == 'ALL STA') {
                map <- us_map[us_map@data$fipsStateCode %in% values$year_state$fipsStateCode, ]
                
                map_data <- values$year_data %>% 
                    filter(declarationTitle == input$declarationTitle)
            } else {
                map <- us_map[us_map@data$fipsStateCode == values$year_state[
                    values$year_state$state_name == input$STATE_NAME, ]$fipsStateCode, 
                ]
                map_data <- values$year_data %>%
                    filter(declarationTitle == input$declarationTitle) %>% 
                    filter(fipsStateCode == values$year_state[
                        values$year_state$state_name == input$STATE_NAME, ]$fipsStateCode
                    )
            }
            map_sta_boder <- aggregate(
                map[, "STATE"], by = list(ID = map@data$STATE), FUN = unique, dissolve = T
            )
            
            map@data <- map@data %>%
                left_join(map_data, by = c('fipsStateCode', 'fipsCountyCode')) 
            
            # information shown on map
            
            map@data$text <- sprintf(
                "<strong>County: %s</strong><br/>federalShareObligated: %s<br/>projectAmount: %s",
                map@data$NAME, map@data$federalShareObligated, map@data$projectAmount
            ) %>% lapply(htmltools::HTML)
            
            colorpal <- colorBin("Blues", map@data$projectAmount, 
                                 na.color = "#f7f7f7", 6, pretty = TRUE)
            
            leaflet(map) %>%
                setMapWidgetStyle(list(background = "#f7f7f7")) %>%
                addPolygons(
                    color = "grey", opacity = 0.2, smoothFactor = 0.3, fillOpacity = 1, weight = 1,
                    fillColor = ~ colorpal(map@data$projectAmount),
                    label = map@data$text
                ) %>%
                addPolylines(data = map_sta_boder, color = "DarkGrey", opacity = 1, weight = 0.8) %>%
                addLegend(
                    pal = colorpal, values = map@data$projectAmount, opacity = 0.9, na.label = "No Hurricane"
                )  
        })
    })
    
    observeEvent(input$declarationTitle, {
        if (input$declarationTitle == 'ALL Hurricane') {
            output$hurricane_sta <- NULL
        }
    })
}

# Run the application 
shinyApp(ui = ui, server = server)
```